---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

Hiven derrigorrezkoa aggregation izena idaztea beheko op eta p bezala
```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?
* The most popular hour of doing order? 
* The most reordered products?
* The less reordered product?
* Plot histogram of number of products in an order.
* What is the most popular department? 

* Question 1*
```{r}
# comando útil 
dbGetQuery(sc, "DESC orders_tbl")
txt1 <- " SELECT order_dow, 
 COUNT(1) AS n_orders_day
FROM orders_tbl
GROUP BY order_dow
ORDER BY   n_orders_day DESC
"
(q1 <- dbGetQuery(sc, txt1))
```
* Question 2*
```{r}
txt2 <- " SELECT order_hour_of_day, 
 COUNT(1) AS n_orders_hour
FROM orders_tbl
GROUP BY order_hour_of_day
ORDER BY n_orders_hour DESC
LIMIT 10
"

(q2 <- dbGetQuery(sc, txt2))
```

* Question 3*
```{r}
txt3 <- " SELECT product_id, 
 COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id, reordered
ORDER BY n_orders DESC
LIMIT 10
"

(q3 <- dbGetQuery(sc, txt3))
```

* Question 4*
```{r}
txt4 <- " SELECT product_id, 
 COUNT(1) AS n_orders
FROM order_products__prior_tbl
WHERE reordered=1
GROUP BY product_id
ORDER BY n_orders
LIMIT 10
"

(q4 <- dbGetQuery(sc, txt4))
```
* Question 5*
```{r}
dbGetQuery(sc, "DESC orders_tbl")
#dist of number of products in each order
txt5 <- "
SELECT order_number,
COUNT(order_id) AS n_products
FROM orders_tbl
GROUP BY order_number
ORDER BY n_products DESC
"

q5 <-
  dbGetQuery(sc, txt5)
q5 %>%
  ggplot(
    aes(reorder(order_number, n_products, function(x) x), 
        n_products)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("order number") +
  ylab("Number of products") +
  theme_minimal()
```


```{r}
dbGetQuery(sc, 
"SELECT n_products
,
COUNT(order_id) AS n_orders
FROM (
      SELECT order_id
      ,
            COUNT(1) AS n_products
      FROM order_products__prior_tbl
      GROUP BY order_id) op
GROUP BY n_products
") %>%
  filter(n_products <= 50) %>%
  ggplot(aes(n_products, n_orders)) +
  geom_col(fill="darkorange") + 
  theme_minimal()
```
```{r}
order_products__prior %>%
  group_by(order_id) %>%
  summarise(n_products = n()) %>%
  group_by(n_products) %>%
  summarise(n_orders = n()) %>%
  filter(n_products <= 50) %>%
  collect %>%
  ggplot(aes(n_products, n_orders)) +
  geom_col(fill="darkorange") + 
  theme_minimal()
```


* Question 6*
```{r}
txt6 <- " SELECT department_id, 
 COUNT(1) AS n_dep
FROM products_tbl
GROUP BY department_id
ORDER BY   n_dep DESC
LIMIT 10
"
(q6 <- dbGetQuery(sc, txt6))
```

## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat  #cat formateatzeko
```
```{r}
data <- list(productId = 24852) # podemos hacer la misma query con distinto producto

product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
* For given department, give a list of 10 most popular product (with templating department).
```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 10) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl
    WHERE department_id = {{depId}}) p
ON op.product_id = p.product_id
"

data <- list(depId = 19)

products_popularity_txt %>%
  whisker.render(data) %>%
  cat  #cat formateatzeko

products_popularity_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .)
```

* For given day, rank department popularity.

```{r, eval = F}
department_popularity_txt <- "
SELECT op.order_id
,   p.product_name
FROM (
    SELECT department_id
    ,   COUNT(1) AS n_department
    FROM products_tbl
    GROUP BY order_id
    ORDER BY n_department DESC
    LIMIT 10) op
LEFT JOIN (
    SELECT order_id
    FROM orders_tbl
    WHERE order_dow = {{dayId}}) p
ON op.order_id = p.order_id
"

data <- list(dayId = 0)

department_popularity_txt %>%
  whisker.render(data) %>%
  cat  #cat formateatzeko

department_popularity_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .)
```

## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```
* Product that is usually orders in numerous baskets

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add product name?


```{r}
"
SELECT apn.product_id
, 
apn.avg_products_number, p.product_name
FROM(
  SELECT pn.product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn 
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100) apn
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON apn.product_id = p.product_id
" %>%
  dbGetQuery(sc, .)
```

* Product that is usually orders in small baskets.

```{r}
"
SELECT apn.product_id
, 
apn.avg_products_number, p.product_name
FROM(
  SELECT pn.product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn 
GROUP BY product_id
ORDER BY avg_products_number 
LIMIT 100) apn
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON apn.product_id = p.product_id
" %>%
  dbGetQuery(sc, .)
```


* For 10 first users and all their orders find days since first order. (CHALLENGING)

```{r}
orders %>% # we are executing sparklyr
  filter(user_id <= 10) %>%
  select(user_id, order_number, days_since_prior_order) %>%
  arrange(user_id, order_number) %>%
  mutate(days_since_first_order = cumsum(days_since_prior_order)) %>%
  mutate(days_since_first_order=ifelse(is.null(days_since_prior_order), 0, days_since_first_order))
```
Now using HIVE SQL (Google-n idatzi hive window functions, )
```{r}
"
SELECT user_id
,   days_since_prior_order
,   
    COALESCE(
        SUM(days_since_prior_order) OVER
        (PARTITION BY user_id ORDER BY order_number ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW), 0) AS
        days_since_first_order
,
    order_number
FROM orders_tbl
WHERE user_id <= 10
ORDER BY user_id
,   order_number
" %>%
  dbGetQuery(sc, .)
```
```{r}
dbGetQuery(sc, "SELECT COUNT(distinct order_number) FROM orders_tbl")
```

